---
title: "Fall 2023 DATA 607 Week 2 Assignment"
author: "Ron Balaban"
date: "2023-09-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 6 Recent popular movies:
- The Barbie Movie
- Oppenheimer
- Spider-Man: Across the Spider-Verse
- The Super Mario Bros. Movie
- Bullet Train
- Indiana Jones and the Dial of Destiny

### Loading the data in
```{r}
library(dplyr)
library(dbplyr)
library(tidyverse)
library(odbc)

# Establishing a connection to MySQL workbench to access the movies table I made there
install.packages("odbc");require("odbc")

con <- dbConnect(odbc::odbc(), 
                 .connection_string = "Driver={MySQL ODBC 8.0 Unicode Driver};", 
                 Server= "localhost", Database= "data", UID= "root", PWD = MySQL_Password,
                 Port= 3306)

# Alternatively, install.packages(c("DBI,"RMySQL")) and con <- dbConnect(RMySQL),host="localhost",user="root",dbname="movies")

moviesdata <- dbGetQuery(con, "SELECT * FROM data.movies")  # Pulling the entire movies table
moviesframe <- as.data.frame(moviesdata) # Convert to dataframe
moviesframe

```

I'd like to know the average score of the movies. However, some movies have not been seen, so we'll ignore the N/A ratings
```{r}

avg_scores <- rowMeans(moviesframe[,3:7],na.rm = TRUE)
# Provides average rating values of the 6 movies across the surveyed participants
# 4.25 3.75 4.20 4.25 3.40 3.00

```

I also considered the approach of having a separate table for those who provided their movie ratings, called 'participants', that stored the following data; Participant_ID,  Participant_Name, and a list of which movies they had seen/rated. I thought to perform a join on movies.rating_X  to  participant.Participant_ID, but the tables were not structured entirely normalized. 

------------------------------------------
# MySQL Workbench table creation

INSERT INTO `data`.`movies`
(`Movie_ID`,
`Movie_Name`,
`Rating_1`,
`Rating_2`,
`Rating_3`,
`Rating_4`,
`Rating_5`)
VALUES
(1,'The Barbie Movie',4,NULL,5,3,5),
(2,'Oppenheimer',4,2,4,5,NULL),
(3,'Spider-Man: Across The Spider-Verse',5,5,4,3,4),
(4,'The Super Mario Bros. Movie',3,4,5,NULL,5),
(5,'Bullet Train',5,3,3,2,4),
(6,'Indiana Jones and The Dial of Destiny',NULL,3,3,4,2)
;



