---
title: "607 Week 10 Text Mining"
author: "Ron Balaban"
date: "2023-11-05"
output: html_document
---


In Text Mining with R, Chapter 2 looks at Sentiment Analysis.  In this assignment, you should start by getting the primary example code from chapter 2 working in an R Markdown document.  You should provide a citation to this base code.  You’re then asked to extend the code in two ways:

- Work with a different corpus of your choosing, and
- Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found through research).

As usual, please submit links to both an .Rmd file posted in your GitHub repository and to your code on rpubs.com.  You make work on a small team on this assignment.





```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Possibly needed Libraries
```{r}
library(httr)  
library(tidytext)
library(readtext)
library(textdata)
library(dplyr)
library(stringr)
library(janeaustenr)
library(tidyr)
library(ggplot2)
library(wordcloud)
library(reshape2)
```


In Text Mining with R, Chapter 2 deals with Sentiment Analysis. (https://www.tidytextmining.com/sentiment.html)
It has 3 sentiment datasets; AFINN, bing, and nrc

```{r}
get_sentiments("afinn")
get_sentiments("bing")
get_sentiments("nrc")
```



Let's perform sentiment analysis on the example cited- books written by Jane Austen 
```{r}
tidy_books <- austen_books() %>%
  group_by(book) %>%
  mutate(
    linenumber = row_number(),
    chapter = cumsum(str_detect(text, 
                                regex("^chapter [\\divxlc]", 
                                      ignore_case = TRUE)))) %>%
  ungroup() %>%
  unnest_tokens(word, text)

#-------------------------------------------------------------------------------

nrc_joy <- get_sentiments("nrc") %>% 
  filter(sentiment == "joy")

#-------------------------------------------------------------------------------
tidy_books %>%
  filter(book == "Emma") %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)

```



Now let's examine the overall sentiment for her books using bing
```{r}
ja_sentiments <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(book, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>% 
  mutate(sentiment = positive - negative)


ja_sentiments
```


And visualize it
```{r}
ggplot(ja_sentiments, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 3, scales = "free_x") +
  ggtitle("Overall Sentiment of Jane Austen Books")
```

Comparing the three sentiment dictionaries
```{r}
pride_prejudice <- tidy_books %>% 
  filter(book == "Pride & Prejudice")

pride_prejudice
```


And visualize them for Pride and Prejudice
```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 80) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")

bing_and_nrc <- bind_rows(
  pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  pride_prejudice %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 80, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)



bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```


```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
  count(sentiment)

#-------------------------------------------------------------------------------
get_sentiments("bing") %>% 
  count(sentiment)
```


Most common positive and negative words
```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

bing_word_counts

#-------------------------------------------------------------------------------
# Visualized

bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)

```

Stop words we ignore
```{r}
custom_stop_words <- bind_rows(tibble(word = c("miss"),  
                                      lexicon = c("custom")), 
                               stop_words)

custom_stop_words
```



Wordcloud to get most common words
```{r}
library(wordcloud)

tidy_books %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100))

#-------------------------------------------------------------------------------
# Reshaped
library(reshape2)

tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("red", "blue"),
                   max.words = 100)


```



Now to perform additional analysis on another topic: Let's use Project gutenbergr which has public books and authors.
(https://github.com/ropensci/gutenbergr)
(https://bookdown.org/Maxine/tidy-text-mining/the-gutenbergr-package.html)


```{r}
devtools::install_github("ropensci/gutenbergr")
```

Package is loaded, can be used
```{r}
library(gutenbergr)

# List of all authors in package: https://www.gutenberg.org/ebooks/
# Lets use Charles Dickens: https://www.gutenberg.org/ebooks/author/37

dickens_books <- gutenberg_works(author == 'Dickens, Charles')

head(dickens_books)

```

Tidying the data
```{r}
tidy_dickens <- dickens_books %>%
  gutenberg_download(meta_fields = 'title') %>%
  group_by(gutenberg_id) %>%
  mutate(linenumber = row_number()) %>%
  ungroup() %>%
  unnest_tokens(word, text)
```



Analyzing sentiment
```{r}
lines_per_index <- 80

dickens_sentiment <- tidy_dickens %>%
  inner_join(get_sentiments('bing'), by = 'word') %>%
  count(title, index = linenumber %/% lines_per_index, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)



dickens_sentiment %>%
  ggplot(aes(index, sentiment, fill = title)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~title, ncol = 10, scales = "free_x")

```


Let's choose just one book and see how the 3 sentiment analyzers from before compare;
```{r}
# A Christmas Carol

a_christmas_carol <- tidy_dickens %>% 
  filter(title == "A Christmas Carol in Prose; Being a Ghost Story of Christmas")



#-------------------------------------------------------------------------------

carol_afinn <- a_christmas_carol %>%
  inner_join(get_sentiments('afinn'), by = 'word') %>%
  group_by(index = linenumber %/% lines_per_index) %>%
  summarize(sentiment = sum(value)) %>%
  mutate(method = "AFINN")

carol_bing_and_nrc <- bind_rows(
  a_christmas_carol %>%
    inner_join(get_sentiments('bing'), by = 'word') %>%
    mutate(method = "Bing et al."),
  a_christmas_carol %>%
    inner_join(get_sentiments('nrc') %>%
                 filter(sentiment %in% c('positive','negative')), by = 'word') %>%
    mutate(method = 'NRC')
) %>%
  count(method, index = linenumber %/% lines_per_index, sentiment) %>%
  pivot_wider(names_from = sentiment, values_from = n, values_fill = 0) %>%
  mutate(sentiment = positive - negative)

bind_rows(carol_afinn, carol_bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) + 
  geom_col(show.legend = FALSE) + 
  facet_wrap(~method, ncol = 1, scales = 'free_y')

```

Most common positive and negative words in our example
```{r}
christmas_carol_word_counts <- a_christmas_carol %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  ungroup()

christmas_carol_word_counts

#-------------------------------------------------------------------------------
# Visualized

christmas_carol_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)




```



